% Term project report -- CSE8003 Learning Automata (VSA inference)
% Compile from this directory: pdflatex term_proj_report.tex
\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xurl}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc,positioning,shapes.geometric,fit,backgrounds}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode}

\lstset{
  language=Python,
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{blue!70!black},
  commentstyle=\color{gray!60},
  stringstyle=\color{green!40!black},
  showstringspaces=false,
  breaklines=true,
  frame=single,
  tabsize=4,
  columns=fullflexible,
  literate={—}{---}1,
}

\title{Variable-Structure Automaton Inference from Interaction Sequences}
\author{Engin Durmaz}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This report documents inference of a Variable-Structure Automaton (VSA) from a
unique dataset of state--action--feedback tuples generated in a P-model
environment. We give the count-based background, methodology for all six course
tasks---dataset fingerprint, estimation, optimal-action identification,
$\varepsilon$-optimality, simulation, perturbation, and a model-assumption
check on per-state penalties---and describe how \texttt{main.py} runs them
end-to-end. Implementation lives under \texttt{../proj\_python/}; code excerpts
are in Appendix~\ref{app:code}.
\end{abstract}

\section{Introduction}

The term project is \textbf{data-driven and reproducible}: from observed
interaction sequences alone we must reconstruct the state--action policy,
transition dynamics, penalty structure, and norm of behavior of an unknown
Variable-Structure Automaton (VSA). The environment is a \textbf{P-model}: binary
feedback $\beta\in\{0,1\}$ encodes favorable vs.\ unfavorable response, and the
assignment assumes that the \emph{global} penalty probability
$c_k = P(\beta{=}1\mid\alpha_k)$ depends on the action $\alpha_k$ but not on the
source state $\phi_i$ (Task~6 tests this assumption empirically).

The system has five states $\{\phi_1,\ldots,\phi_5\}$, three actions
$\{\alpha_1,\alpha_2,\alpha_3\}$, and many sequences indexed by
\texttt{sequence\_id}. Each row of the CSV is one step:
\[
  (\text{sequence\_id},\,\text{step},\,\Phi_{\mathrm{src}}(n),\,\Phi_{\mathrm{dst}}(n),\,\alpha(n),\,\beta(n)).
\]
Before modeling we report a \textbf{dataset fingerprint} (row count, number of
sequences, mean length, empirical penalty rate per action, checksum). The six
tasks then proceed as follows:
\begin{enumerate}[label=\textbf{Task \arabic*:}, leftmargin=*]
  \item \textbf{Estimation} --- empirical counts
        $N(\phi_i,\alpha_k)$, $N(\phi_i,\alpha_k,\phi_j)$, $N(\alpha_k,\beta)$;
        policy $P(\alpha_k\mid\phi_i)$, transitions $P(\phi_j\mid\phi_i,\alpha_k)$,
        and penalties $c_k$.
  \item \textbf{Optimal action} --- $\alpha^\ast \in \arg\min_k c_k$.
  \item \textbf{$\varepsilon$-optimality} ---
        $\varepsilon = 1 - \min_i P(\alpha^\ast\mid\phi_i)$ over visited states.
  \item \textbf{Simulation} --- synthetic trajectories from the learned model;
        compare marginals and penalty rates to the real CSV.
  \item \textbf{Perturbation} --- remove a random fraction of rows, refit
        Tasks~1--4, and assess stability of $c_k$, policy, and $\varepsilon$.
  \item \textbf{Model assumption check} --- estimate
        $c_k(\phi_i)=P(\beta{=}1\mid\alpha_k,\phi_i)$ per state and compare across
        $\phi_i$ (and, under perturbation, across trials) to judge the P-model.
\end{enumerate}

All tasks are implemented in \texttt{../proj\_python/}. The orchestrator
\texttt{main.run\_all} prints the fingerprint, runs Tasks~1--4 on the full
dataset (baseline), optionally repeats Tasks~1--4 on row-subsampled copies when
\texttt{perturbation\_rate}${}>0$, summarizes Task~5 metrics, and prints combined
Task~6 tables once at the end (baseline plus every perturbation trial when
applicable). Each task module can also be run standalone and writes outputs under
\texttt{outputs/}. This report aligns notation with the course assignment and
with that code; Section~\ref{sec:background} gives VSA context and empirical
estimators (Section~\ref{sec:estimators}); Section~\ref{sec:methodology} maps
each task to code.

\section{Background}\label{sec:background}

\subsection{Variable-structure automata and the P-model}\label{sec:vsa}

A \textbf{Variable-Structure Automaton} (VSA) is a stochastic decision model in
which an agent visits discrete states $\phi_i$, chooses actions $\alpha_k$, and
receives environment feedback. In this project the state set is
$\{\phi_1,\ldots,\phi_5\}$, the action set is $\{\alpha_1,\alpha_2,\alpha_3\}$,
and each interaction step is recorded in the dataset together with a sequence
identifier and step index. The learning goal is to recover, from many such
sequences only, four objects required by the assignment:
\begin{itemize}[noitemsep]
  \item a \textbf{state--action policy} $P(\alpha_k\mid\phi_i)$,
  \item \textbf{transition dynamics} $P(\phi_j\mid\phi_i,\alpha_k)$,
  \item an \textbf{environmental penalty structure} (probabilities of unfavorable
        feedback), and
  \item a \textbf{norm of behavior} (e.g.\ how often the optimal action is taken
        in each state, via $\varepsilon$-optimality).
\end{itemize}

The environment is a \textbf{P-model}: feedback $\beta\in\{0,1\}$ indicates a
favorable ($0$) or unfavorable ($1$) response. The working assumption is that
the \emph{global} penalty $c_k=P(\beta{=}1\mid\alpha_k)$ depends on the action
but not explicitly on the source state---so one penalty vector suffices for
Tasks~1--5. Task~6 examines whether that assumption is consistent with the data
when penalties are estimated separately in each state.

\subsection{Approach used in this project}\label{sec:approach}

The implementation is \textbf{purely data-driven}: every quantity is an empirical
frequency or a ratio of counts from the provided CSV, followed by row-wise
normalization where conditional distributions are required. We do not fit
parametric models, run gradient-based learning, or impose structure beyond what
the counts support. Consistency checks (policy rows and transition rows sum to
one; global $c_k$ matches the sample mean of $\beta$ given $\alpha$) validate
the arithmetic.

The six assignment tasks form a pipeline: a \textbf{dataset fingerprint}
(Section~\ref{sec:estimators}) validates scale and raw penalty rates; \textbf{estimation}
builds counts and conditional probabilities; \textbf{optimal-action} and
\textbf{$\varepsilon$-optimality} analyses use the policy and global penalties;
\textbf{simulation} draws synthetic trajectories from the learned model and
compares marginals to the real data; \textbf{perturbation} randomly deletes rows
and refits to test robustness; \textbf{model critique} compares per-state
penalties $c_k(\phi_i)$ across $\phi_i$ to assess the P-model. Section~\ref{sec:methodology}
maps each step to code; Appendix~\ref{app:code} lists the corresponding excerpts.

\subsection{Empirical estimators}\label{sec:estimators}

All formulas below are evaluated on a (possibly subsampled) interaction table.
Each CSV row is one step; we write $n\in R$ for row indices and $S$ for the set
of distinct \texttt{sequence\_id} values. For a fixed sequence $s\in S$,
$J_s$ is how many rows belong to $s$ (that sequence's length in steps).

\paragraph{Dataset fingerprint.}
Before modeling, the assignment requires five summaries of the raw table:
\begin{itemize}[noitemsep]
  \item \textbf{Total number of rows} $|R|$ (one per step in the file).
  \item \textbf{Number of sequences} $|S|$ (distinct \texttt{sequence\_id}).
  \item \textbf{Mean sequence length}
        $\displaystyle \bar{L}=\frac{1}{|S|}\sum_{s\in S} J_s$
        (average of the per-sequence lengths $J_s$).
  \item \textbf{Empirical penalty rate per action} $\alpha_k$:
        fraction of rows with $\alpha(n)=\alpha_k$ that have $\beta(n)=1$,
        \[
          \frac{1}{\#\{n\in R : \alpha(n)=\alpha_k\}}
          \sum_{\substack{n\in R \\ \alpha(n)=\alpha_k}} \beta(n).
        \]
  \item \textbf{Checksum}
        $\displaystyle C=\sum_{n\in R} \texttt{step}(n)$.
\end{itemize}
On the full dataset, each per-action rate above equals the global penalty
$c_k$ used in estimation (both are the sample mean of $\beta$ given $\alpha_k$).

\paragraph{Raw counts.}
The assignment requires these three tables \emph{before} any probabilities are
reported: they are the empirical basis for the policy $P(\alpha_k\mid\phi_i)$,
the transitions $P(\phi_j\mid\phi_i,\alpha_k)$, and the penalties $c_k$.
Each entry counts how many rows $n\in R$ match a pattern on
\texttt{phi\_src}, \texttt{phi\_dst}, \texttt{alpha}, and \texttt{beta}
(written $\Phi_{\mathrm{src}}(n)$, $\Phi_{\mathrm{dst}}(n)$, $\alpha(n)$,
$\beta(n)$ below).

\medskip
\noindent\textbf{$N(\phi_i,\alpha_k)$ --- actions taken in a state.}
\begin{equation}
  N(\phi_i,\alpha_k)
  = \#\{\, n\in R : \Phi_{\mathrm{src}}(n)=\phi_i,\; \alpha(n)=\alpha_k \,\}.
\end{equation}

\medskip
\noindent\textbf{$N(\phi_i,\alpha_k,\phi_j)$ --- transitions under an action.}
\begin{equation}
  N(\phi_i,\alpha_k,\phi_j)
  = \#\{\, n\in R : \Phi_{\mathrm{src}}(n)=\phi_i,\;
  \alpha(n)=\alpha_k,\; \Phi_{\mathrm{dst}}(n)=\phi_j \,\}.
\end{equation}

\medskip
\noindent\textbf{$N(\alpha_k,\beta)$ --- environment feedback per action.}
\begin{equation}
  N(\alpha_k,\beta)
  = \#\{\, n\in R : \alpha(n)=\alpha_k,\; \beta(n)=\beta \,\}.
\end{equation}
\noindent Write $r_i=\sum_k N(\phi_i,\alpha_k)$ for the row total over actions in
state $\phi_i$.

\paragraph{Conditional estimates (Tasks 1--4).}
For visited states ($r_i>0$),
\begin{equation}
  P(\alpha_k\mid\phi_i) = \frac{N(\phi_i,\alpha_k)}{r_i}.
\end{equation}
For each action $\alpha_k$, transition probabilities normalize each row of
$N(\phi_i,\alpha_k,\phi_j)$ over $\phi_j$ (empty rows yield zeros). The
\textbf{global} P-model penalty is
\begin{equation}
  c_k = P(\beta{=}1\mid\alpha_k)
  = \frac{N(\alpha_k,1)}{N(\alpha_k,0)+N(\alpha_k,1)}.
\end{equation}
Task~2 sets $\alpha^\ast\in\arg\min_k c_k$. Task~3 uses visited states
$\mathcal{V}=\{\phi_i : r_i>0\}$ and
$\varepsilon = 1 - \min_{\phi_i\in\mathcal{V}} P(\alpha^\ast\mid\phi_i)$.
Task~4 samples $\alpha\sim P(\cdot\mid\phi)$, $\phi'\sim P(\cdot\mid\phi,\alpha)$,
and $\beta\sim\mathrm{Bernoulli}(c_\alpha)$ along each real sequence length.

\paragraph{Per-state penalties (Task 6).}
To test action-only penalties, for each pair $(\phi_i,\alpha_k)$ with
$N(\phi_i,\alpha_k)>0$,
\begin{equation}
  c_k(\phi_i) = P(\beta{=}1\mid\alpha_k,\phi_i)
  = \frac{N(\phi_i,\alpha_k,\beta{=}1)}{N(\phi_i,\alpha_k)},
\end{equation}
where $N(\phi_i,\alpha_k,\beta{=}1)$ counts penalized steps with that state and
action. Under a strict P-model, $c_k(\phi_i)$ should vary little in $\phi_i$ for
fixed $k$; a diagnostic is $\max_i c_k(\phi_i)-\min_i c_k(\phi_i)$ per action.

\section{Methodology}\label{sec:methodology}

All runnable code lives in \texttt{../proj\_python/}. Two execution modes exist:
\textbf{orchestrated} (\texttt{python main.py} $\rightarrow$ \texttt{load\_all}
$\rightarrow$ \texttt{run\_all}) and \textbf{standalone} (each \texttt{run\_*}
module with \texttt{output\_dir} under \texttt{outputs/}). Formulas are in
Section~\ref{sec:estimators}; Appendix~\ref{app:code} lists the matching source.

\subsection{General flow}\label{sec:procedure}

The end-to-end pipeline is implemented in \texttt{main.run\_all} (Appendix~\ref{app:main}).
It always validates the raw CSV, fits the VSA quantities on the full dataset, and
optionally repeats the fit on random row subsamples to test stability. Per-state
penalty tables are printed once at the end so the console log stays readable.
Figure~\ref{fig:flow} summarizes \texttt{run\_all} with decisions and loops;
Algorithm~\ref{alg:flow} states the same logic in pseudocode. A standalone run executes only
\texttt{RunTasks1To4} for a single module at a time, reading and writing CSVs under
\texttt{outputs/}.

\begin{algorithm}[htbp]
  \caption{General flow (\texttt{main.run\_all})}
  \label{alg:flow}
  \begin{algorithmic}[1]
    \Require CSV path, \texttt{perturbation\_rate}, \texttt{experiment\_count},
             \texttt{random\_seed}
    \State $df_{\mathrm{full}} \gets \texttt{load\_vsa\_dataset}(\text{path})$
           \Comment{\texttt{util.py}}
    \State \texttt{run\_fingerprint}(\text{path})
           \Comment{$|R|$, $|S|$, $\bar{L}$, penalty rates, $C$}
    \State $baseline \gets \textsc{RunTasks1To4}(df_{\mathrm{full}},\ \text{baseline})$
    \State $results \gets [\,baseline\,]$
    \If{\texttt{perturbation\_rate} $> 0$}
      \For{$t = 1$ \textbf{to} \texttt{experiment\_count}}
        \State $df_p \gets \texttt{\_drop\_rows\_randomly}(df_{\mathrm{full}},\ \text{rate},\ rng)$
        \State append $\textsc{RunTasks1To4}(df_p,\ t)$ to $results$
      \EndFor
      \State \texttt{\_print\_experiments\_summary}($baseline$, perturbed runs)
    \EndIf
    \State \texttt{\_print\_c\_k\_per\_state\_runs}($results$)
    \Statex
    \Function{RunTasks1To4}{$df$, $label$}
      \State $est \gets \texttt{run\_estimation}(df,\ \texttt{print\_penalty\_per\_state=False})$
      \State $opt \gets \texttt{run\_optimal\_action}(est)$
      \State $eps \gets \texttt{run\_epsilon\_optimality}(est,\ opt)$
      \State $sim \gets \texttt{run\_simulation}(est,\ \text{path})$
      \State \Return \texttt{PipelineRunResult}($label$, $est$, $opt$, $eps$, $sim$)
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{figure}[htbp]
  \centering
  \resizebox{0.92\linewidth}{!}{%
  \begin{tikzpicture}[
    font=\scriptsize,
    >=Stealth,
    node distance=4.5mm,
    startstop/.style={draw, rounded corners=2pt, minimum width=2.4cm,
      minimum height=6mm, align=center, fill=gray!12},
    proc/.style={draw, rounded corners=2pt, minimum width=3.1cm,
      minimum height=6mm, align=center, fill=blue!8},
    decision/.style={draw, diamond, aspect=2.4, inner sep=0.5pt,
      align=center, fill=orange!14},
    taskchain/.style={draw, rounded corners=2pt, minimum width=3.4cm,
      minimum height=8mm, align=center, fill=blue!5, font=\tiny},
    line/.style={thick},
  ]
    \node[startstop] (start) {Start};
    \node[proc, below=of start] (load) {Load interaction dataset};
    \node[proc, below=of load] (fp) {Dataset fingerprint\\(Task~0)};
    \node[taskchain, below=of fp] (t14) {Tasks 1--4 on full data\\(estimate $\rightarrow$ $\alpha^\ast$
      $\rightarrow$ $\varepsilon$ $\rightarrow$ simulate)};
    \node[proc, below=of t14] (store) {Store baseline run};
    \node[decision, below=of store] (dpert) {Perturbation\\rate $> 0$?};

    \node[proc, below left=10mm and 14mm of dpert] (t6only)
      {Print per-state $c_k(\phi_i)$\\(Task~6, baseline only)};
    \node[startstop, below=of t6only] (endL) {Return results};

    \node[proc, below right=10mm and 8mm of dpert] (drop)
      {Randomly drop rows\\(fraction $p$)};
    \node[taskchain, below=of drop] (t14p)
      {Tasks 1--4 on subsample};
    \node[proc, below=of t14p] (append) {Append trial to results};
    \node[decision, below=of append] (dmore) {More trials?\\$t < T$};
    \node[proc, below=of dmore] (t5) {Print perturbation summary\\(Task~5)};
    \node[proc, below=of t5] (t6all)
      {Print per-state $c_k(\phi_i)$\\(Task~6, all runs)};
    \node[startstop, below=of t6all] (endR) {Return results};

    \begin{scope}[on background layer]
      \node[draw, dashed, rounded corners=3pt, inner sep=5pt, fit=(drop)(t14p)(append)(dmore),
        label={[font=\tiny]above:{$t = 1,\ldots,T$}}] (loopbox) {};
    \end{scope}

    \draw[line] (start) -- (load);
    \draw[line] (load) -- (fp);
    \draw[line] (fp) -- (t14);
    \draw[line] (t14) -- (store);
    \draw[line] (store) -- (dpert);
    \draw[line] (dpert) -- node[above left, font=\tiny] {no} (t6only);
    \draw[line] (t6only) -- (endL);
    \draw[line] (dpert) -- node[above right, font=\tiny] {yes} (drop);
    \draw[line] (drop) -- (t14p);
    \draw[line] (t14p) -- (append);
    \draw[line] (append) -- (dmore);
    \draw[line] (dmore) -- node[right, font=\tiny] {no} (t5);
    \draw[line] (t5) -- (t6all);
    \draw[line] (t6all) -- (endR);
    \draw[line] (dmore.west) -- ++(-0.55,0) |- node[pos=0.22, left, font=\tiny] {yes} (drop.west);
  \end{tikzpicture}%
  }
  \caption{Flowchart for \texttt{run\_all}: baseline Tasks~1--4 on the full dataset,
    optional perturbation loop over $T$ trials (random row removal, refit, append),
    then Task~5 aggregate reporting when perturbation is enabled and Task~6
    per-state penalty tables (baseline only if $p=0$, else all stored runs).}
  \label{fig:flow}
\end{figure}

\subsection{Implementation details}\label{sec:implementation}
\sloppy

The subsubsections below describe how each module realizes one part of the
methodology. For each file we state the goal, point to the matching appendix
listing, and list the main third-party libraries used (the standard library
supplies \texttt{pathlib}, \texttt{dataclasses}, \texttt{argparse}, and
\texttt{typing} where needed).

\subsubsection{Software map}\label{sec:software-map}

Table~\ref{tab:software-map} lists each project module, its main entry point(s),
return type, and the matching appendix listing (file by file, no task labels).

\begin{table}[H]
  \centering
  \caption{Source modules in \texttt{../proj\_python/}. An orchestrated run
    (\texttt{python main.py}) uses every row via \texttt{main.run\_all}; standalone
    runs call one \texttt{run\_*} function at a time and pass CSVs through
    \texttt{outputs/}.}
  \label{tab:software-map}
  \footnotesize
  \setlength{\tabcolsep}{5pt}
  \begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash}p{4.75cm}
    >{\raggedright\arraybackslash}X
    >{\raggedright\arraybackslash}p{2.35cm}
    >{\centering\arraybackslash}p{0.75cm}@{}}
    \toprule
    \textbf{Module} & \textbf{Entry point(s)} & \textbf{Return type} & \textbf{Ref.} \\
    \midrule
    \texttt{fingerprint.py}
      & \texttt{run\_fingerprint}
      & \texttt{Fingerprint\-Result}
      & \ref{app:fingerprint} \\
    \texttt{estimation.py}
      & \texttt{run\_estimation}
      & \texttt{Estimation\-Result}
      & \ref{app:estimation} \\
    \texttt{optimal\_action\_}\allowbreak\texttt{identification.py}
      & \texttt{run\_optimal\_action}
      & \texttt{Optimal\-Action\-Result}
      & \ref{app:optimal} \\
    \texttt{e\_optimality\_analysis.py}
      & \texttt{run\_epsilon\_optimality}
      & \texttt{Epsilon\-Optimality\-Result}
      & \ref{app:epsilon} \\
    \texttt{simulation.py}
      & \texttt{run\_simulation}
      & \texttt{Simulation\-Result}
      & \ref{app:simulation} \\
    \texttt{main.py}
      & \texttt{run\_all}, \texttt{load\_all}
      & \texttt{Pipeline\-Run\-Result} (list)
      & \ref{app:main} \\
    \texttt{util.py}
      & \texttt{load\_vsa\_dataset}
      & \texttt{DataFrame}
      & \ref{app:util} \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsubsection{Dataset fingerprint (\texttt{fingerprint.py})}\label{sec:impl-fingerprint}

\textbf{Goal.}\; Before modeling, report $|R|$, $|S|$, $\bar{L}$, empirical penalty
rate per action, and checksum $C$ (Section~\ref{sec:estimators}) so the submission
can be validated against the assignment fingerprint.

\textbf{Code.}\;
Appendix~\ref{app:fingerprint} (\texttt{run\_fingerprint}, lines~86--99), invoked at
the start of Algorithm~\ref{alg:flow}.

\textbf{Libraries.}\; \textbf{pandas} (\texttt{groupby}, \texttt{mean}, \texttt{sum}).

\subsubsection{Estimation (\texttt{estimation.py})}\label{sec:impl-estimation}

\textbf{Goal.}\; From the interaction table, build raw count matrices and derive
$P(\alpha_k\mid\phi_i)$, $P(\phi_j\mid\phi_i,\alpha_k)$, global $c_k$, and
per-state $c_k(\phi_i)$; run consistency checks (Section~\ref{sec:estimators}).

\textbf{Code.}\;
Appendix~\ref{app:estimation} (\texttt{run\_estimation}, lines~104--189).
From \texttt{main}, per-state penalties are computed but not printed
(\nolinkurl{print_penalty_per_state=False}).
Standalone \nolinkurl{estimation.py} writes CSVs under
\nolinkurl{outputs/task1_estimation/}.

\textbf{Libraries.}\; \textbf{pandas} (\texttt{groupby}, \texttt{unstack}, normalization);
\textbf{NumPy} (\texttt{allclose} for checks).

\subsubsection{Optimal action (\texttt{optimal\_action\_identification.py})}\label{sec:impl-optimal}

\textbf{Goal.}\; Identify $\alpha^\ast \in \arg\min_k c_k$ from the penalty vector and
record ties deterministically.

\textbf{Code.}\;
Appendix~\ref{app:optimal} (\texttt{run\_optimal\_action}, lines~109--128); uses
\texttt{estimation.penalty} in memory or \texttt{penalty.csv} when run standalone.

\textbf{Libraries.}\; \textbf{pandas} (sorting, framing $c_k$); \textbf{NumPy}
(\texttt{isclose} for tie detection).

\subsubsection{$\varepsilon$-optimality (\texttt{e\_optimality\_analysis.py})}\label{sec:impl-epsilon}

\textbf{Goal.}\; Compute
$\varepsilon = 1 - \min_{\phi_i \in \mathcal{V}} P(\alpha^\ast\mid\phi_i)$ over
visited states only.

\textbf{Code.}\; Appendix~\ref{app:epsilon}
(\texttt{\_compute\_core} / \texttt{run\_epsilon\_optimality}, lines~110--130).

\textbf{Libraries.}\; \textbf{pandas} (policy and count matrices); \textbf{NumPy}
(\texttt{min}, \texttt{isclose}).

\subsubsection{Simulation (\texttt{simulation.py})}\label{sec:impl-simulation}

\textbf{Goal.}\; Generate synthetic trajectories with the learned policy, transitions,
and Bernoulli penalties; compare data vs.\ simulated marginals and penalty rates.

\textbf{Code.}\; Appendix~\ref{app:simulation}
(\texttt{\_simulate\_block}, \texttt{run\_simulation}, lines~134--184).

\textbf{Libraries.}\; \textbf{pandas} (marginals, alignment); \textbf{NumPy}
(\texttt{default\_rng}, sampling, binomial draws).

\subsubsection{Perturbation experiment (\texttt{main.py})}\label{sec:impl-perturbation}

\textbf{Goal.}\; Delete a random fraction of rows, refit the model on each subsample,
and compare $\alpha^\ast$, $\varepsilon$, and simulation L1 to the baseline
(Algorithm~\ref{alg:flow}, perturbation branch).

\textbf{Code.}\; Appendix~\ref{app:main}:
\texttt{\_drop\_rows\_randomly} (lines~36--51),
\texttt{run\_all} perturbation loop (lines~319--344),
\texttt{\_print\_experiments\_summary} (lines~162--255).

\textbf{Libraries.}\; \textbf{NumPy} (\texttt{default\_rng}, \texttt{choice} for row
indices); \textbf{pandas} (\texttt{drop}, \texttt{reset\_index}).

\subsubsection{Model assumption check (\texttt{estimation.py}, \texttt{main.py})}\label{sec:impl-task6}

\textbf{Goal.}\; Form $c_k(\phi_i)$ for each state--action pair and compare across
states (spread diagnostic); optionally across perturbation trials
(Section~\ref{sec:estimators}).

\textbf{Code.}\; \texttt{penalty\_per\_state} in Appendix~\ref{app:estimation};
printing and spread helpers in Appendix~\ref{app:main}
(\texttt{\_spread\_across\_states}, \texttt{\_print\_c\_k\_per\_state\_runs},
lines~70--78, 150--159).

\textbf{Libraries.}\; \textbf{pandas} (per-state table); \textbf{NumPy} (max/min spread,
absolute differences).
Also see \texttt{\_run\_tasks\_1\_to\_4} and \texttt{run\_all} in
Appendix~\ref{app:main} (lines~91--133, 302--346).

\subsubsection{Shared data loading (\texttt{util.py})}\label{sec:impl-util}

\textbf{Goal.}\; Load the headerless interaction CSV into a typed
\texttt{pandas.DataFrame} and define shared result containers passed between modules.

\textbf{Code.}\; Appendix~\ref{app:util}: column constants,
\texttt{EstimationResult}, \texttt{OptimalActionResult}, and
\texttt{load\_vsa\_dataset}.

\textbf{Libraries.}\; \textbf{pandas} (\texttt{read\_csv}, frame types).

\fussy

\section{Experiments}

\textbf{Data.} The provided file
\nolinkurl{../Engin-vsa_dataset_full.csv} (one row per step; Greek state/action
labels) is the default input for \texttt{estimation.py}.

\textbf{Reproducibility.} From \texttt{../proj\_python/}:
\begin{enumerate}
  \item Task~1 only (writes CSVs and \texttt{report.txt}):
        \nolinkurl{python estimation.py}
  \item Task~2 after Task~1 files exist:
        \nolinkurl{python optimal_action_identification.py}
  \item Task~3 after Tasks~1--2 outputs exist:
        \nolinkurl{python e_optimality_analysis.py}
  \item Task~4 after Task~1 (and real CSV for marginals):
        \nolinkurl{python simulation.py}
  \item End-to-end (\texttt{main.py}, \texttt{load\_all}):
        \nolinkurl{python main.py}.\\
        Defaults: full CSV;
        \nolinkurl{perturbation_rate=20},
        \nolinkurl{random_seed=42}.
\end{enumerate}
Console logs from \texttt{python main.py} include the fingerprint, full Task~1--4
reports per run, Task~5 aggregates, and combined Task~6 per-state penalty tables.

\section{Conclusion}

We summarized VSA background, a count-based approach, and estimators in
Section~\ref{sec:background}. The pipeline recovers policy, transitions, and
global penalties (Task~1), identifies $\alpha^\ast$ (Task~2), quantifies
$\varepsilon$-optimality over visited states (Task~3), and validates a
simulator against data marginals (Task~4). Task~5 stress-tests robustness under
random row removal; Task~6 compares $c_k(\phi_i)$ across states to assess the
P-model. Together, these six tasks satisfy the assignment structure from
fingerprint through model critique.

\appendix
\section[Supplementary source code]{Supplementary source code listing}\label{app:code}

Excerpts below are taken from \texttt{../proj\_python/}, one subsection per file.
Line ranges match the repository at report time; together they implement the
flow in Section~\ref{sec:procedure} (Algorithm~\ref{alg:flow}) and the estimators in
Section~\ref{sec:estimators}.

\subsection{\texttt{fingerprint.py}}\label{app:fingerprint}

\lstinputlisting[firstline=86,lastline=99]{../proj_python/fingerprint.py}

\subsection{\texttt{estimation.py}}\label{app:estimation}

\lstinputlisting[firstline=104,lastline=189]{../proj_python/estimation.py}

\subsection{\texttt{optimal\_action\_identification.py}}\label{app:optimal}

\lstinputlisting[firstline=109,lastline=128]{../proj_python/optimal_action_identification.py}

\subsection{\texttt{e\_optimality\_analysis.py}}\label{app:epsilon}

\lstinputlisting[firstline=110,lastline=130]{../proj_python/e_optimality_analysis.py}

\subsection{\texttt{simulation.py}}\label{app:simulation}

\lstinputlisting[firstline=134,lastline=184]{../proj_python/simulation.py}

\subsection{\texttt{main.py}}\label{app:main}

\texttt{\_drop\_rows\_randomly}:
\lstinputlisting[firstline=36,lastline=51]{../proj_python/main.py}

\texttt{\_run\_tasks\_1\_to\_4}:
\lstinputlisting[firstline=91,lastline=133]{../proj_python/main.py}

\texttt{\_print\_c\_k\_per\_state\_runs}:
\lstinputlisting[firstline=150,lastline=159]{../proj_python/main.py}

\texttt{\_print\_experiments\_summary} (spread and table-delta helpers):
\lstinputlisting[firstline=70,lastline=78]{../proj_python/main.py}
\lstinputlisting[firstline=214,lastline=255]{../proj_python/main.py}

\texttt{run\_all}:
\lstinputlisting[firstline=302,lastline=346]{../proj_python/main.py}

\subsection{\texttt{util.py}}\label{app:util}

Column constants, shared result types, and CSV loading:
\lstinputlisting[firstline=21,lastline=54]{../proj_python/util.py}
\lstinputlisting[firstline=57,lastline=60]{../proj_python/util.py}
\lstinputlisting[firstline=62,lastline=66]{../proj_python/util.py}
\lstinputlisting[firstline=69,lastline=82]{../proj_python/util.py}

\end{document}
